### 2016-02-29 paper-1/main.tex
Stat and Prob Letter draft started.

\begin{figure}[ht]
\includegraphics[width=\textwidth]{KanePlot.pdf}
\caption{
A hypothetical realization of the breast cancer trial.
}
\label{fig:kane_viz}
\end{figure}

@@

Let $\,b_1, b_2, \ldots \,$ denote a sequence of independent, identically 
distributed, Bernoulli random variables with $\mathbb{E}I_{\{b_i=1\}}=p$ and
$\mathbb{E}I_{\{b_i = 0\}} = 1-p$, for 
probability parameter $0\leq p \leq 1$. In the clinical trial setting 
$\,b_i = 1$ corresponds to a successful outcome.  Let $s$ and $t$ be 
positive integers.  Define the stopped negative binomial (SNB) random 
variable $Y$ as the smallest 
integer value such that $\,\{b_1, \ldots , b_Y\}\,$ contains {\em either} 
$\,s\,$ successes {\em or} $\,t\,$ failures. That is, either 
$\sum_i^Y b_i s$ or else $\sum_i^Y 1-b_i = t$.

The distribution of $\,Y\,$ has support on integer values in the range 
\begin{equation*}               
     \min(s,t) \leq \; Y \;\leq s+t-1  \label{range.y.eq}
\end{equation*}
and it is distributed as
\begin{equation} \label{eqn:pmf}
\mathbb{E}\  I_{\{Y=k\}} = S(k, p, s) I_{\{s \leq k\}} + 
  S(k, 1-p, t) I_{\{ t \leq k \}}
\end{equation}
where $I_{\{f\}}$ is one if $f$ is true and zero otherwise and
\begin{equation} \label{eqn:N}
S(k, p, s) = {k-1 \choose s-1} p^s (1-p)^{k-s} 
\end{equation}
$S$ is the negative binomial probability 
In~\ref{eqn:pmf} we can see that the mass at 
any point on the support of the distribution may come from $S(p)$; $S(1-p)$;
or both of their probabilities.

@@

To show the result in Eqn.~\ref{eqn:pmf} consider the 
process $\mathbf{X} = \left\{X(k) : k = 0,1,... \right\}$ 
with $X(0)=0$ and
\begin{equation*} \label{eqn:proc}
X_{k+1} = X_k + b_{k+1} I_{\{ k-t < X_k < s\}}.
\end{equation*}
The process can be conceptualized as a series of coin 
flips that stops when either $s$ successes or $t$ failures are reached. 
At each step a coin is flipped. If it is heads, the process advances 
one diagonally in the 
positive horizontal and vertical direction in Fig.~\ref{fig:kane_viz}. 
Otherwise, it advances in the positive horizontal direction only. When the 
process stops either $X_k = s$ or $X_k = k-t$.

@@

Let $\mathcal{I}_p(s, t)$ be the {\em regularized incomplete beta function}, which is 
also one minus the c.d.f. of the negative binomial distribution. This
function satisfies $\mathcal{I}_p(s, t) = 1-\mathcal{I}_{1-p}(t, s)$.  
Then it can be proven
that \ref{eqn:pmf} is a valid probability mass by showing that the
probability mass function sums to one.
\begin{align*}
R = \sum_{i=0}^{t-1} &{i+s-1 \choose i} p^s (1-p)^i +
\sum_{j=0}^{s-1}  {j+t-1 \choose j} p^j  (1-p)^t \\
   &= 1-\mathcal{I}_p(s, t) + 1 - \mathcal{I}_{1-p}(t, s) \\
   &= 1. 
\end{align*}
\end{proof}

@@

Let us assume that $p$ has a Beta distribution, with constant prior 
parameters $\alpha$ and $\beta$. Then the closed form 
posterior distribution of the SNB is
\begin{align}
f \left(k | \mathbf{b}, \alpha, \beta \right) &= 
  {k-1 \choose s-1} \frac{B\left(\alpha+s, k-s+\beta \right)}{B(\alpha, \beta)} 
  I_{\{s \leq k \leq s+k-1\}} + \nonumber \\
  & {k-1 \choose k-t} 
  \frac{B\left(\alpha + k - t, t+\beta\right)}{B(\alpha, \beta)} 
  I_{\{t \leq k \leq s+k-1\}}
\end{align}
where $B$ is the Beta function and $\mathbf{b}$ is the realization 
of Bernoulli data.

@@

\item[\bibref{Kane 2015}]
Kane MJ. (2015). snb: The Stopped Negative Binomial. R package version 0.1. 2015. Available from \url{https://github.com/kaneplusplus/snb}

@@

@@

\item[\bibref{Simon 1989}]
Simon R.  Optimal two-stage designs for Phase II clinical trials. {\it Controlled Clinical Trials\/} 1989; {\bf 10}: 1--10.

\item[\bibref{Uppuluri 1970}] 
Uppuluri VRR, Blot WJ. ``A probability distribution arising in a riff-shuffle.'' {\it Random Counts  in Scientific Work, 1: Random Counts in Models and Structures}, 1970. G.P. Patil (editor), University Park: Pennsylvania State University Press, pp  23--46.

\item[\bibref{Zhang 2000}]
Zhang Z, Burtness BA, Zelterman D.  The maximum negative binomial distribution. {\em Journal of Statistical Planning and Inference\/}  2000; {\bf 87}: 1--19.

>>>>

\begin{figure}[t!]
\includegraphics[width=\textwidth]{KanePlot.pdf}
\caption{
A hypothetical realization of the breast cancer trial.
}
\label{fig:kane_viz}
\end{figure}

@@

Let $\,b_1, b_2, \ldots \,$ denote a sequence of independent, identically 
distributed, Bernoulli random variables with $\mathbb{P}[b_i=1]=p$ and
$\mathbb{P}[b_i = 0] = 1-p$, for 
probability parameter $0\leq p \leq 1$. In the clinical trial setting 
$\,b_i = 1$ corresponds to a successful outcome.  Let $s$ and $t$ be 
positive integers.  Define the stopped negative binomial (SNB) random 
variable $Y$ as the smallest 
integer value such that $\,\{b_1, \ldots , b_Y\}\,$ contains {\em either} 
$\,s\,$ successes {\em or} $\,t\,$ failures. That is, either 
$\sum_i^Y b_i = s$ or $\sum_i^Y 1-b_i = t$.

The distribution of $\,Y\,$ has support on integer values in the range 
\begin{equation*}               
     \min(s,t) \leq \; Y \;\leq s+t-1  \label{range.y.eq}
\end{equation*}
and it is distributed as
\begin{equation} \label{eqn:pmf}
\mathbb{E}\  I_{\{Y=k\}} = S(k, p, s) \ I_{\{s \leq k\}} + 
  S(k, 1-p, t) \ I_{\{ t \leq k \}}
\end{equation}
where $I_{\{f\}}$ is one if $f$ is true and zero otherwise and
\begin{equation} \label{eqn:N}
S(k, p, s) = {k-1 \choose s-1} p^s (1-p)^{k-s} 
\end{equation}
$S$ is the negative binomial probability 
In~\ref{eqn:pmf} we can see that the mass at 
any point on the support of the distribution may come from $S(p)$; $S(1-p)$;
or both of their probabilities.

@@

To show the result in Eqn.~\ref{eqn:pmf} consider the 
process $\mathbf{X} = \left\{X(k) : k = 0,1,... \right\}$ 
with $X(0)=0$ and
\begin{equation*} \label{eqn:proc}
X_{k+1} = X_k + b_{k+1} \ I_{\{ k-t < X_k < s\}}.
\end{equation*}
The process can be conceptualized as a series of coin 
flips that stops when either $s$ successes or $t$ failures are reached. 
At each step a coin is flipped. If it is heads, the process advances 
one diagonally in the 
positive horizontal and vertical direction in Fig.~\ref{fig:kane_viz}. 
Otherwise, it advances in the positive horizontal direction only. When the 
process stops either $X_k = s$ or $X_k = k-t$.

@@

Let $\mathcal{I}_p(s, t)$ be the {\em regularized incomplete beta function}, which is 
also one minus the c.d.f. of the negative binomial distribution. This
function satisfies $\mathcal{I}_p(s, t) = 1-\mathcal{I}_{1-p}(t, s)$
[\ref{abramowitz1964}].
Then it can be proven
that \ref{eqn:pmf} is a valid probability mass by showing that the
probability mass function sums to one.
\begin{align*}
R = \sum_{i=0}^{t-1} &{i+s-1 \choose i} p^s (1-p)^i +
\sum_{j=0}^{s-1}  {j+t-1 \choose j} p^j  (1-p)^t \\
   &= 1-\mathcal{I}_p(s, t) + 1 - \mathcal{I}_{1-p}(t, s) \\
   &= 1. 
\end{align*}
\end{proof}

@@

Let us assume that $p$ has a Beta distribution, with constant prior 
parameters $\alpha$ and $\beta$. Then the closed form 
posterior distribution of the SNB is
\begin{align}
f \left(k | \mathbf{b}, \alpha, \beta \right) &= 
  {k-1 \choose s-1} \frac{B\left(\alpha+s, k-s+\beta \right)}{B(\alpha, \beta)} 
  \ I_{\{s \leq k \leq s+k-1\}} + \nonumber \\
  & {k-1 \choose k-t} 
  \frac{B\left(\alpha + k - t, t+\beta\right)}{B(\alpha, \beta)} 
  \ I_{\{t \leq k \leq s+k-1\}}
\end{align}
where $B$ is the Beta function and $\mathbf{b}$ is the realization 
of Bernoulli data.

@@

\item[\bibref{Kane 2015}]
MJ Kane. (2015). snb: The Stopped Negative Binomial. R package version 0.1. 2015. Available from \url{https://github.com/kaneplusplus/snb}

@@

\item[\bibref{abramowitz1964}]
M Abramowitz, IA Stegun. {\it Handbook of mathematical functions: with 
formulas, graphs, and mathematical tables}. Courier Corporation. 1964;
{\bf 55}: 263.

@@

\item[\bibref{Simon 1989}]
R Simon.  Optimal two-stage designs for Phase II clinical trials. {\it Controlled Clinical Trials\/} 1989; {\bf 10}: 1--10.

\item[\bibref{Uppuluri 1970}] 
VRR Uppuluri, WJ Blot. ``A probability distribution arising in a riff-shuffle.'' {\it Random Counts  in Scientific Work, 1: Random Counts in Models and Structures}, 1970. G.P. Patil (editor), University Park: Pennsylvania State University Press, pp  23--46.

\item[\bibref{Zhang 2000}]
Z Zhang, BA Burtness, D Zelterman.  The maximum negative binomial distribution. {\em Journal of Statistical Planning and Inference\/}  2000; {\bf 87}: 1--19.
